---
title: "p8105_hw2_as6445"
author: "Ayako Sekiya"
date: "2022-09-27"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

#Problem 1=

This question is not graded. Solutions will be posted later. Make sure to include the code here still.

I will first import the csv file and janior::clean names to clean the data. 

Then, I selected the variables that I wanted to include in my data set. 

I then changed the character variables for the entry variable into logical variables. 
```{r import}
transit_data = read_csv(file = "./Data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
janitor::clean_names() %>%
select(transit_data, line:station_location, entrance_location, entry, vending, ada) %>%
recode(entry,  )
```

The following dataset includes the following variables: Line, "Station Name", "Station Latitude", "Station Longitude", "Station Location", "Entrance Location", Entry, Vending, ADA. 

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?

# Problem 2

I am importing the first sheet of the spreadsheet which includes information from trashwheel using `sheet`. I am also restricting the columns that are imported into the dataframe to exclude columns with notes. I used `drop_na()` to 

```{r tidy data}
trashwheel_data = 
  readxl::read_excel("./Data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet=1, skip= 1, range=cell_cols("A:N")) %>%
  na.omit() %>%
  as.integer(round("Sports Balls", digits=1))
  
professorwheel_data = 
  readxl::read_excel("./Data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet=2, skip= 1, range=cell_cols("A:N")) %>%
  na.omit() %>%
  as.integer(round("Sports Balls", digits=0))
```

# Problem 3

```{r}
pols-month = read_csv(file = "./Data/pols-month.csv")
unemployment = read_csv(file = "./Data/unemployment.csv")
snp = read_csv(file = "./Data/snp.csv")
```

